\documentclass[12pt]{svmono}

\input{preamble}

\usepackage{graphicx}

\usepackage[agsm]{harvard}

\title{\textbf{\sc Deliberative Autonomy for Marine Robotics}}
\author{}
\date{\small Monterey Bay Aquarium Research Institute \\
7700 Sandholdt Rd.\\
Moss Landing,  CA, 95039\\
{\tt }}

\begin{document}

\maketitle

\begin{abstract}
\\
  We describe a general purpose Artificial Intelligence based control
  architecture that implements the sense-plan-act paradigm. The
  Teleo-Reactive EXecutive (T-REX) framework is deliberates about
  future states, plans for actions and executes activities while
  monitoring plans for anomalous conditions. It is unusual in being
  operational and in routine use for operational oceanography,
  providing a new tool for science to sample and observe the dynamic
  coastal ocean. Plans are no longer scripted a priori but synthesized
  onboard with high-level directives instead of low level
  commands. Further, the architecture uses multiple control loops for
  a 'divide and conquer' problem solving strategy allowing for
  incremental computational model building, robust and focused failure
  recovery, ease of software development and ability to use legacy or
  non-native computational paradigms. Vehicle adaptation and sampling
  occurs in situ with additional modules which can be selectively used
  depending on the application in focus. Abstraction in problem
  solving allows different applications to be programmed relatively
  easily, with little to no changes to the core search engine thereby
  making software engineering sustainable. The representational
  ability to deal with time and resources coupled with Machine
  Learning techniques for event detection, allows balancing shorter
  term benefits with longer term needs, an important need as AUV
  hardware becomes more robust allowing persistent ocean sampling and
  observation.
  
%  % Traditional approaches to commanding marine robots have relied on
%  %  pre-scripted plans with a priori defined waypoints. These have
%  %  served the oceanographic community well in enabling robots to
%  %  cost-effectively perform surveys traditionally undertaken by
%  %  ship-board observations. However, in dynamic coastal environments
%  %  where is substantial interaction between atmospheric, oceanographic,
%  %  estuarine/riverine and land-sea interaction processes requires a
%  %  robot to be adaptive and receptive to opportunistic
%  %  science. Sustained use of such autonomous platforms in military
%  %  applications have also pushed.

\end{abstract}

\input{intro.tex}

\input{related.tex}

\input{arch.tex}

\input{concepts.tex}

\input{results.tex}

\input{future.tex}

\input{conclude.tex}

\section{Acknowledgements}

The authors are funded by a block grant from the David and Lucile
Packard Foundation to MBARI. 

\bibliography{references}

\end{document}
